Towards the Development of an Intelligent Building Energy Management System -- iBEMS \thanks{The results presented in this paper were produced by a group of senior undergraduate students: Brian Lauer and Elliot Watkins }

### 2021-07-19 paperIEEE-IECON2021.tex
Paper submitted to IEEE IECON 2021

\begin{IEEEkeywords}
  Building energy management, internet of things, multi-agent systems, energy monitoring  
\end{IEEEkeywords}

@@

% The previous senior project group worked on creating a remote motor control system with a Raspberry Pi and XBee module. To supply power directly to the motor and XBee with a single power supply, they used a buck-boost converter capable of stepping down power supply voltage to 3.3 V. They implemented a Python script that uses the Linux command \texttt{nmap} to search for the Raspberry Pi on the network. The Raspberry Pi connects to the transmitter XBee via a serial connection to transmit commands to the receiver XBee mounted to the motor. They were able to simply toggle the motor on and off and partially implemented this functionality in BEMOSS (Building Energy Management Open Source Software). A second portion of their project was implementing an HVAC control algorithm capable of controlling the temperature of a home with the Linear Quadratic Regulator control algorithm.

% During a research project led by Miah \textit{et al.,} the DC motor interface was successfully
% implemented into the BEMOSS platform which required adding a great deal of source code
% to the platform to fully implement the device as BEMOSS does not support adding
% new devices currently. The research also consisted of adding new
% features such as a logo to the
% platform for the motor. A possible improvement our platform will make is helping
% developers on the project add new devices more easily.

% One of the improvements our project makes over the previous is the use of UDP (user datagram protocol) multicasting which does not require any superuser privileges to locate an
% embedded computer (Beaglebone Blue, for example). This implementation is also more convenient
% because it does not require installing any packages as multicasting is natively
% supported by most embedded Linux systems. If in the future, multiple embedded
% computers are needed to be discovered, the multicasting system works elegantly
% as each device can simply listen on a provided multicast group which consists of
% an IP address and port number.

@@

At the center of the iBEMS architecture is the iBEMS core. This core interacts
with multiple peripherals including IoT devices and a weather API over the
internet. In this work, we developed the iBEMS to support for two 
IoT devices displayed in the high-level architecture shown in Fig.~\ref{fig:highLevelArchitecture}. %
%
\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.15]{figs/overallConnectionDiagram.pdf}
  \caption{High-level architecture of the proposed iBEMS.}
  \label{fig:highLevelArchitecture}
\end{figure}
%
These devices are the WeMo Insight Switch and an embedded computer known as the Beaglebone
Blue.  The core is also able to
connect the internet to download weather data for a local city through
OpenWeatherMap. Fig.~\ref{fig:functional_bd} gives the lower level functional block diagram of the system
architecture. %
%
\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.15]{figs/functionalBlockDiagram.pdf}
  \caption{Functional Block Diagram}
  \label{fig:functional_bd}
\end{figure}
%
The inputs include the user input through the web server, the device status
updates (On/Off, power usage, etc...) and the weather data. The iBEMS Core
itself consists of a Flask Web Server, 2 databases (Apache Cassandra and
SQLite), and 3 software agents (Discovery, Control, and Scheduling).

@@

Fig.~\ref{fig:systemComponentInterconnection} provides an explanation of how
the components of the system interact with each other. %
%
\begin{figure}
  \centering
  \includegraphics[scale=0.15]{figs/overallDiagram.pdf}
  \caption{Interconnection between system components}
  \label{fig:systemComponentInterconnection}
\end{figure}
%
The Apache Cassandra database and SQLite database are accessible from the web server, control agent, discovery agent, and scheduling agent. The web server queries the time series database and metadata database for vastly different purposes. Power plotting requires querying the Apache Cassandra database. To distinguish between the different devices on the front end, the metadata database is used. The agents utilize the Apache Cassandra database for storing and extracting time-series data. The agents use the metadata database to distinguish between different devices.

@@

More specifically, the publish subscribe model of communication is used where
each agent communicates via a socket where each socket responds to a specific
topic. The topics and port numbers are listed below along with their
corresponding agents in Table~\ref{tab:pubsubspecs}. %
%
\begin{table}[htbp]
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        Agent & Topic & Port Number\\
        \hline
        Discovery Agent & "discovery" & 5556\\
        Scheduling Agent & "scheduling" & 5556\\
        Control Agent & "control" & 5556\\
        \hline
    \end{tabular}
    \caption{Publish/Subscribe Communication Specifications}
    \label{tab:pubsubspecs}
\end{table}
The agents contained in the software are associated with a Python class of the
same name. Each class is instantiated and run with specifications declared in
Table~\ref{tab:pubsubspecs}.

The flow chart of method \texttt{setDeviceStatus} defined in the \texttt{ControlAgent} class is shown in
Figure~\ref{fig:setDeviceStatus}. The purpose of this method is to set the
status of the desired device. When the method is called, the data dictionary is
obtained which contains the device ID. With this device ID, the API type can be
determined using the SQLite database. Subsequently, the appropriate device API
module can be imported. If the power state exists as a key in the device data
dictionary, the power state of the WeMo Insight Switch will be set as ON or OFF.
If the duty cycle key exists, the duty cycle of the single board computer motor
drivers is set. If the shutdown key is set, the single board computer is
shutdown. In order to reconnect to the single board computer, the board must be
started and receiver program ran. If necessary, the device must be reconnected
to the WiFi network. %

\begin{figure}[htbp]
    \centering
    \includegraphics[scale=0.2]{figs/agents/setDeviceStatus.pdf}
    \caption{setDeviceStatus Flow Chart}
    \label{fig:setDeviceStatus}
\end{figure}

To obtain necessary data quantities from the device, the
\texttt{getDeviceStatus} method defined in \texttt{ControlAgent}, defined in Figure~\ref{fig:getDeviceStatus}, is called. Once the proper data parameters are
obtained, the \texttt{getState} function defined in the devices API module can
be called with the data parameters such as power state and
status. Once the agent has finished sending these to the device, a request is
sent to the web server to inform the server that it completed searching for the
required device data. %
%
\begin{figure}[htbp]
    \centering
    \includegraphics[scale=0.3]{figs/agents/getDeviceStatus.pdf}
    \caption{getDeviceStatus Flow Chart}
    \label{fig:getDeviceStatus}
\end{figure}

Every device connected to iBEMS will have a thread associated with it started from the control agent. This thread will run the \texttt{periodicQueryBehavior} method defined in \texttt{ControlAgent} to periodically query for device data and store in the corresponding Cassandra database table. A user specified delay is placed in the class's module to prevent the thread from overwhelming the system.

@@

The scheduling agent is able to update the device status of each device dependent on whether a schedule is set for the specific time period and day. One of the methods that is called, \texttt{updateScheduleFromServer} is invoked when the user presses the \say{Update Schedule} button from the UI on the scheduling page. A flow chart of this method is shown in Figure~\ref{fig:updateScheduleFromServer}. When this method is called the desired schedule is loaded in as a dictionary and parsed to extract the necessary periods and given day. First the times are stored in military time in the database for more convenient storage and parsing later on. Then, a table is created in the Apache Cassandra database where the device ID and day are embedded in the name. The previous table's data is destroyed to ensure there are no errors regarding overlapping schedules. Lastly, all the scheduling periods extracted from the dictionary are stored as rows in the table. 
\begin{figure}[htbp]
    \centering
    \includegraphics[scale=0.2]{figs/agents/updateScheduleFromServer.pdf}
    \caption{updateScheduleFromServer Flow Chart}
    \label{fig:updateScheduleFromServer}
\end{figure}
A second method that is run inside each device thread is \texttt{periodicDeviceUpdateMethod}, defined in Figure~\ref{fig:periodicDeviceUpdateMethod}, which is started when the discovery agent sets the discovered device to active. By first loading the device's API name, the scheduling agent can begin checking for scheduling periods inside an infinite loop. The status dictionary is queried from the Cassandra database and loaded into a dictionary. All these scheduling periods are looped over for the current day while constantly checking whether the current time in seconds is within the starting and ending time. A conversion of both starting and ending time to seconds is necessary as it is challenging to work with these quantities in standard and military time when comparing times. If the current time does lie inside the scheduling period, the device's state is set dependent on the API. 
\begin{figure}[htbp]
    \centering
    \includegraphics[scale=0.2]{figs/agents/periodicDeviceUpdateMethod.pdf}
    \caption{periodicDeviceUpdateMethod Flow Chart}
    \label{fig:periodicDeviceUpdateMethod}
\end{figure}

>>>>

\begin{IEEEkeywords}
  Building energy management, embedded computer, internet of things, multi-agent systems, smart plug,  web server
\end{IEEEkeywords}

@@

@@

At the center of the iBEMS architecture is the iBEMS core. This core interacts
with multiple peripherals including IoT devices and a weather API over the
internet. In this work, we developed the iBEMS to support for two 
IoT devices displayed in the high-level architecture shown in Fig.~\ref{fig:highLevelArchitecture}. %
%
\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.15]{figs/overallConnectionDiagram.pdf}
  \caption{High-level architecture of the proposed iBEMS.}
  \label{fig:highLevelArchitecture}
\end{figure}
%
These devices are the WeMo Insight Switch and an embedded computer known as the Beaglebone
Blue.  The core is also able to
connect the internet to download weather data for a local city through
OpenWeatherMap. Fig.~\ref{fig:functional_bd} gives the lower level functional block diagram of the system
architecture. %
%
\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{figs/functionalBlockDiagram.pdf}
  \caption{Functional Block Diagram}
  \label{fig:functional_bd}
\end{figure}
%
The inputs include the user input through the web server, the device status
updates (On/Off, power usage, etc...) and the weather data. The iBEMS Core
itself consists of a Flask Web Server, 2 databases (Apache Cassandra and
SQLite), and 3 software agents (Discovery, Control, and Scheduling).

@@

Fig.~\ref{fig:systemComponentInterconnection} provides an explanation of how
the components of the system interact with each other. %
%
\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figs/overallDiagram.pdf}
  \caption{Interconnection between system components}
  \label{fig:systemComponentInterconnection}
\end{figure}
%
The Apache Cassandra database and SQLite database are accessible from the web server, control agent, discovery agent, and scheduling agent. The web server queries the time series database and metadata database for vastly different purposes. Power plotting requires querying the Apache Cassandra database. To distinguish between the different devices on the front end, the metadata database is used. The agents utilize the Apache Cassandra database for storing and extracting time-series data. The agents use the metadata database to distinguish between different devices.

@@

More specifically, the publish subscribe model of communication is used where
each agent communicates via a socket where each socket responds to a specific
topic. The topics and port numbers are listed below along with their
corresponding agents in Table~\ref{tab:pubsubspecs}. %
%
\begin{table}[htbp]
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        Agent & Topic & Port Number\\
        \hline
        Discovery Agent & "discovery" & 5556\\
        Scheduling Agent & "scheduling" & 5556\\
        Control Agent & "control" & 5556\\
        \hline
    \end{tabular}
    \caption{Publish/Subscribe Communication Specifications}
    \label{tab:pubsubspecs}
\end{table}
%
The agents contained in the software are associated with a Python class of the
same name. Each class is instantiated and run with specifications declared in
Table~\ref{tab:pubsubspecs}.

The flow chart of method \texttt{setDeviceStatus} defined in the \texttt{ControlAgent} class is shown in
Figure~\ref{fig:setDeviceStatus}. The purpose of this method is to set the
status of the desired device. When the method is called, the data dictionary is
obtained which contains the device ID. With this device ID, the API type can be
determined using the SQLite database. Subsequently, the appropriate device API
module can be imported. If the power state exists as a key in the device data
dictionary, the power state of the WeMo Insight Switch will be set as ON or OFF.
If the duty cycle key exists, the duty cycle of the single board computer motor
drivers is set. If the shutdown key is set, the single board computer is
shutdown. In order to reconnect to the single board computer, the board must be
started and receiver program ran. If necessary, the device must be reconnected
to the WiFi network. %
%
\begin{figure}[htbp]
    \centering
    \includegraphics[scale=0.2]{figs/agents/setDeviceStatus.pdf}
    \caption{setDeviceStatus Flow Chart}
    \label{fig:setDeviceStatus}
\end{figure}
%
To obtain necessary data quantities from the device, the
\texttt{getDeviceStatus} method defined in \texttt{ControlAgent}, defined in Figure~\ref{fig:getDeviceStatus}, is called. Once the proper data parameters are
obtained, the \texttt{getState} function defined in the devices API module can
be called with the data parameters such as power state and
status. Once the agent has finished sending these to the device, a request is
sent to the web server to inform the server that it completed searching for the
required device data. %
%
\begin{figure}[htbp]
    \centering
    \includegraphics[scale=0.3]{figs/agents/getDeviceStatus.pdf}
    \caption{getDeviceStatus Flow Chart}
    \label{fig:getDeviceStatus}
\end{figure}
%
Every device connected to iBEMS will have a thread associated with it started from the control agent. This thread will run the \texttt{periodicQueryBehavior} method defined in \texttt{ControlAgent} to periodically query for device data and store in the corresponding Cassandra database table. A user specified delay is placed in the class's module to prevent the thread from overwhelming the system.

@@

The scheduling agent is able to update the device status of each device dependent on whether a schedule is set for the specific time period and day. One of the methods that is called, \texttt{updateScheduleFromServer} is invoked when the user presses the \say{Update Schedule} button from the UI on the scheduling page. A flow chart of this method is shown in Figure~\ref{fig:updateScheduleFromServer}. When this method is called the desired schedule is loaded in as a dictionary and parsed to extract the necessary periods and given day. First the times are stored in military time in the database for more convenient storage and parsing later on. Then, a table is created in the Apache Cassandra database where the device ID and day are embedded in the name. The previous table's data is destroyed to ensure there are no errors regarding overlapping schedules. Lastly, all the scheduling periods extracted from the dictionary are stored as rows in the table. %
%
\begin{figure}[htbp]
    \centering
    \includegraphics[scale=0.3]{figs/agents/updateScheduleFromServer.pdf}
    \caption{updateScheduleFromServer Flow Chart}
    \label{fig:updateScheduleFromServer}
\end{figure}
%
A second method that is run inside each device thread is \texttt{periodicDeviceUpdateMethod}, defined in Figure~\ref{fig:periodicDeviceUpdateMethod}, which is started when the discovery agent sets the discovered device to active. By first loading the device's API name, the scheduling agent can begin checking for scheduling periods inside an infinite loop. The status dictionary is queried from the Cassandra database and loaded into a dictionary. All these scheduling periods are looped over for the current day while constantly checking whether the current time in seconds is within the starting and ending time. A conversion of both starting and ending time to seconds is necessary as it is challenging to work with these quantities in standard and military time when comparing times. If the current time does lie inside the scheduling period, the device's state is set dependent on the API. %
%
\begin{figure}[htbp]
    \centering
    \includegraphics[scale=0.2]{figs/agents/periodicDeviceUpdateMethod.pdf}
    \caption{periodicDeviceUpdateMethod Flow Chart}
    \label{fig:periodicDeviceUpdateMethod}
\end{figure}
%
